Decentralized state machine using Nakamoto (probabilistic) consensus\\\medskip Research Project 2022

### 2022-11-02 article/article.tex
typos

In Raft, each node can be either Leader, Follower or Candidate. In regular functioning, one node will be the leader and all the others will follow.
The Leader node sends regular heartbeat requests to notify Followers that it is still running. If they stop receiving the heartbeat, they switch to the Candidate status.

@@

Servers are represented by nodes. A server which received the request to add the log is considered "colored".

@@

The term in the sums is considered 0 if it does not make sense.
First sum represents a step of the resolution of one group of requests, coming from a semi-colored node, with tc (ts) the number of colored (semi colored) nodes in the sample.
Second sum represents a step of the resolution of one group of requests, coming from a colored node.
To do the calculation, we assume that requests from semi-colored nodes resolve faster.
\begin{multline*}
    P(c, s, rc, rs) = \\ \sum_{tv=0}^{k+1}\sum_{ts=0}^{k+1} {n-c-s+tv \choose tv}{s-tv-1+ts \choose ts}{c-ts \choose k-tv-ts}P(c-ts, s-tv+ts, rc-ts, rs+1-tv)
    \\ + \sum_{ts=0}^{k+1} {n-c-s+rs \choose tv}{s-rs+ts \choose ts}{c-ts-1 \choose k-rs-ts}P(c-ts, s-rs+ts, rc+1-ts, 0)
\end{multline*}
\\
The starting point of the Markov process is 1 semi-colored node and 1 group of requests from a semi-colored node. 
\begin{equation*}
    P(0, 1, 0, 1) = 1
\end{equation*}
\\
With these formulas, we can compute the probability of not reaching consensus : see Appendices.
The scripts which were used to compute the graphs in Appendices are available at \href{https://github.com/Nicolascrd/researchProjectConsensus}{github.com/Nicolascrd/researchProjectConsensus}. They use Dynamic Programming.

@@

\paragraph{Raft (classical)}
When the leader is up and giving instruction to all the followers, followers don't have to initiate requests.
Therefore in normal functioning, the leader node will make $n-1$ requests every heartbeat.
In addition to this, every client request will trigger $n-1$ requests from the leader, to update followers. If the request was directed from the client to a follower, one request is initiated
by the follower to transfer the request to the leader.

@@

\paragraph{Fail-stop in Snowball}
If a node crashs in Snowball, some add-log request will not find a majority and therefore the consensus will be slower and require more requests.
However, in production, the Snowball state machine would include a mechanism to update the network knowledge among members, which can take the form of a Snowball run, and therefore continue as if no node had failed.
Snowball is also very efficient at managing crash failures.

@@

We demonstrated that it was possible to build a decentralized state machine, that could be based on a Nakamoto consensus algorithm and had interesting properties. 
In normal running state, the number of requests to reach consensus with Snowball is a function of $k * \beta$ for each node. In Raft, the leader must handle $n-1$ request on its own. 
Raft has one consensus for the whole system, whereas in Snowball to each value correspond a different consensus. For most applications, classical consensus is probably the way to go. Still, it would be useful to lead some more research on the topic. 
In particular, it would be interesting to approximate the value of ε $\forall n, k, \beta$. 
Also, a comparison focused on Byzantine Fault Tolerance would also be interesting. To do that one would need to use PBFT \cite{pbft} instead of the Raft.

>>>>

In Raft, each node can be either Leader, Follower or Candidate. In regular functioning, one node will be the leader and all the others will follow.
The Leader sends regular heartbeat requests to notify Followers that it is still running. If they stop receiving the heartbeat, they switch to the Candidate status, in which they ask other nodes to vote for them in order to become Leader.

@@

Servers are represented by nodes. A server which received the request to add the log is considered "colored", otherwise it is considered "neutral".

@@

The term in the sums is considered 0 if it does not make sense.
The first sum represents a step of the resolution of one group of requests, coming from a semi-colored node, with tc (ts) the number of colored (semi colored) nodes in the sample.
The second sum represents a step of the resolution of one group of requests, coming from a colored node.
To do the calculation, we assume that requests from semi-colored nodes resolve faster.
\begin{multline*}
    P(c, s, rc, rs) = \\ \sum_{tv=0}^{k+1}\sum_{ts=0}^{k+1} {n-c-s+tv \choose tv}{s-tv-1+ts \choose ts}{c-ts \choose k-tv-ts}P(c-ts, s-tv+ts, rc-ts, rs+1-tv)
    \\ + \sum_{ts=0}^{k+1} {n-c-s+rs \choose tv}{s-rs+ts \choose ts}{c-ts-1 \choose k-rs-ts}P(c-ts, s-rs+ts, rc+1-ts, 0)
\end{multline*}
\\
The starting point of the Markov process is 1 semi-colored node and 1 group of requests from a semi-colored node. 
\begin{equation*}
    P(0, 1, 0, 1) = 1
\end{equation*}
\\
With these formulas, we can compute the probability of not reaching consensus : see Appendices.
The scripts which were used to compute the graphs in Appendices are available at \href{https://github.com/Nicolascrd/researchProjectConsensus}{github.com/Nicolascrd/researchProjectConsensus}. They use Dynamic Programming.

@@

\paragraph{Raft (classical)}
When the leader is up and giving instructions to all the followers, followers don't have to initiate requests.
Therefore in normal functioning, the leader node will make $n-1$ requests every heartbeat.
In addition to this, every client request will trigger $n-1$ requests from the leader, to update followers. If the request was directed from the client to a follower, one request is initiated
by the follower to transfer the request to the leader.

@@

\paragraph{Fail-stop in Snowball}
With no system to update knowledge among participating nodes, nodes crashes can force nodes to use more requests to reach consensus. 
Indeed, if in the sample there are enough crashed nodes to prevent the success of the request, the node will have to start a new group of requests. That requires a significant number of nodes to crash nonetheless.
In production, the Snowball state machine would include a mechanism to update the network knowledge among members, which can take the form of a Snowball run, and therefore continue as if no node had failed.
Snowball would then be very efficient at managing crash failures.

@@

We demonstrated that it was possible to build a decentralized state machine, that could be based on a Nakamoto consensus algorithm and had interesting properties. 
In normal running state, the number of requests to reach consensus with Snowball is a function of $k * \beta$ for each node. In Raft, the leader must handle $n-1$ request on its own. 
Raft has one consensus for the whole system, whereas in Snowball to each value corresponds a different consensus run. For most applications, classical consensus is probably the way to go. Still, it would be useful to lead some more research on the topic. 
In particular, it would be interesting to approximate the value of ε $\forall n, k, \beta$. 
A comparison focused on Byzantine Fault Tolerance would also be interesting. To do that one would need to use PBFT \cite{pbft} instead of the Raft.
